---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bigmds

<!-- badges: start -->
<!-- badges: end -->

*MDS* is a statistic tool for reduction of dimensionality, using as input a distance matrix of dimensions *n × n*. 
When *n* is large, classical algorithms suffer from computational problems and *MDS* configuration can not be obtained.

With this package, we address these problems by means of three algorithms: 

- Divide-and-conquer MDS.
- Fast MDS.
- Interpolation MDS.

The main idea of these methods is based on partitioning the dataset into small pieces, where classical methods can work.
*Fast MDS* was developed by *Yang, T., J. Liu, L. McMillan, and W. Wang (2006)*, whereas 
*divide-and-conquer MDS* and *interpolation MDS* were developed by *Delicado P. and C. Pachón-García (2021).*

To obtain more information, please read this [paper](https://arxiv.org/abs/2007.11919).

## Installation

You can install directly from CRAN with:

``` r
# install.packages("bigmds")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pachoning/bigmds")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
set.seed(42)
library(bigmds)
x <- matrix(data = rnorm(4*10000), nrow = 10000) %*% diag(c(9, 4, 1, 1))

divide_mds_conf <- divide_conquer_mds(x = x, l = 200, c_points = 2*2, r = 2, n_cores = 1, dist_fn = stats::dist)
head(divide_mds_conf$points)
divide_mds_conf$eigen
divide_mds_conf$GOF

fast_mds_conf <- fast_mds(x = x, l = 200, s_points = 2*2, r = 2, n_cores = 1, dist_fn = stats::dist)
head(fast_mds_conf$points)
fast_mds_conf$eigen
fast_mds_conf$GOF

interpolation_mds_conf <- interpolation_mds(x = x, l = 200, r = 2, n_cores = 1, dist_fn = stats::dist)
head(interpolation_mds_conf$points)
interpolation_mds_conf$eigen
interpolation_mds_conf$GOF
```

With the implementation of *classical MDS*, it takes much more time to obtain a MDS configuration due to computational 
problems. Try it yourself!

```{r, example2, eval=FALSE}
x <- matrix(data = rnorm(4*10000, sd = 10), nrow = 10000)
dist_matrix <- stats::dist(x = x)
mds_result <- stats::cmdscale(d = dist_matrix, k = 2, eig = TRUE)
```
