---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bigmds

<!-- badges: start -->
<!-- badges: end -->

We present a set of algorithms for *Multidimensional Scaling (MDS)* to be used with large datasets. *MDS* is a statistic 
tool for reduction of dimensionality, using as input a distance matrix of dimensions *n × n*. When *n* is large, 
classical algorithms suffer from computational problems and *MDS* configuration can not be obtained.

With this package, we address these problems by means of three algorithms: 

- Divide and Conquer MDS
- Fast MDS
- MDS based on Gower interpolation

The main idea of these methods is based on partitioning the dataset into small pieces, where classical methods can work.
*Fast MDS* was developed by *Yang, Tynia & Liu, Jinze & Mcmillan, Leonard & Wang, Wei. (2006)*, whereas 
*Divide and Conquer MDS* and *MDS based on Gower interpolation* were developed by *Delicado and Pachon-Garcia, 2020.*

To obtain more information, please read this [paper](https://arxiv.org/abs/2007.11919).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pachoning/bigmds")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(bigmds)
x <- matrix(data = rnorm(4*10000, sd = 10), nrow = 10000)

dc_mds <- divide_conquer_mds(x = x, l = 100, s = 8, k = 2)
head(dc_mds$points)
dc_mds$eigen

f_mds <- fast_mds(x = x, l = 100, s = 8, k = 2)
head(f_mds$points)
f_mds$eigen

g_mds <- gower_interpolation_mds(x = x, l = 100, k = 2)
head(g_mds$points)
g_mds$eigen
```

With the implementation of *classical MDS*, it is not possible to obtain a MDS configuration due to computational 
problems. Try it yourself!

```{r, example2, eval=FALSE}
library(bigmds)
x <- matrix(data = rnorm(4*10000, sd = 10), nrow = 10000)
dist_matrix <- stats::dist(x = x)
mds_result <- stats::cmdscale(d = dist_matrix, k = 2, eig = TRUE)
```
